---
title: "convert a genetic map to recombination map"
geometry: margin=1.5cm
output: pdf_document
date: 2022-07-29
---

| chr_phy | phy_pos | chr_gen | gen_pos |
|---|---|---|---|
| physical chr| physical position (in base)| genetic chr | genetic position |

# 1. Load data (to edit if rerun)
```{r,echo=T}
data = read.table('~/Domisol_thesis/rawdata_system/maize/carte_maize.txt',h=T)
data$phy_pos = data$coordinate # i just rename this column to match the column syntax 
head(data)
```
# 2. Exemple for one chromosome

```{r,echo=T}
data$phy_pos = data$phy_pos / 1e6 # transform base position in Mb position

smooth_strengh = 0.7 # choose a value between 0 and +infinite.
#The higher is the value, the smoother the map will be

sub_data = subset(data,subset = chr_phy == 1)
sm = smooth.spline(x=sub_data$phy_pos,y=sub_data$gen_pos,spar=smooth_strengh)

### check if the smooth is enough
predx = seq(0.1,max(sub_data$phy_pos),by=0.1)
predy = predict(sm,x=predx,deriv=0)$y
plot(sub_data$phy_pos,sub_data$gen_pos,type='p')
lines(predx,predy,col='red')
#### you must obtain a growing monotonous dot line , and the red line must be in the dot line.

pred_deriv = predict(sm,x=predx,deriv=1)$y
plot(predx,pred_deriv,type='l',xlab='position (Mb)',ylab='rec_rate (cM/Mb)')
# here the map seems smooth enough to remove all the noise

sub_rec_map = data.frame('chr'=1,'start' = predx - (0.1/2 - 1e-6) , 
						 'end' = predx + 0.1/2 , 'rec_rate' = pred_deriv) 
head(sub_rec_map)
```
The map generated express the recombination rate in cM/Mb. The recombination rate is express
in probability of recombination per base, per generation, per individual.
Furthermore, the physical position (start and end) are express in Mb too, which doesn’t match 
the file format use in RIDGE (which is in pb)

| chr | start (bp) | end (bp) | rec_rate |
|-----|------------|----------|----------|
| chr name | start position | end position of window | recombination rate |

# 3. Transform cM/Mb to recombination rate using haldane transformation

\[ d_{AB} = - 50.ln(1-2.r_{AB}) \]
with $d_{AB}$ the genetic distance between A and B in cM. Our map give us a result in cM,
so
\[ \frac{cM}{Mb}.Mb=d_{AB} \]
And $r_{AB}$ is the recombination rate between A and B. So $r_{AB}$ is equal to 
\[ r_{AB}=\frac{1-e^{d_{AB}/-50}}{2}\]


```{r}
sub_rec_map$rec_rate = sub_rec_map$rec_rate * 0.1 / 1e5 #to transform into d_AB / locus 
#size to set the distance between d_AB at 1 base
sub_rec_map$rec_rate = (1-exp(sub_rec_map$rec_rate * 100/-50))/2
sub_rec_map$start = sub_rec_map$start * 1e6
sub_rec_map$end = sub_rec_map$end * 1e6
head(sub_rec_map)
print(mean(sub_rec_map$rec_rate))
```
